### 2017-01-26 manuscript.rmd
reorganized code in chunks and functions to improve clarity and make additional subanalyses easier

	# These clade dates are from the KMRR code, they got them from http://timetree.org/
	calibration_times = data.frame(
		age.min =
			c( 20, 92, 167, 8, 42, 74, 296, 535, 104, 937, 29, 722, 441, 65, 15, 1215, 
				414, 371, 162, 25, 74, 77, 86 ), 
		clade = 
			c("Hominoidea", "Euarchontoglires", "Mammalia", "Homininae", 
					"Simiiformes", "Primates", "Amniota", "Vertebrata", "Eutheria", 
					"Bilateria", "Catarrhini", "Chordata", "Euteleostomi", 
					"Haplorrhini", "Hominidae", "Opisthokonta", "Sarcopterygii", 
					"Tetrapoda", "Theria", "Murinae", "Sciurognathi", "Rodentia", "Glires"
			),
		stringsAsFactors=FALSE
	)

	# Functions

@@

	#' Adjust branch lengths of a phylogenetic tree to make it ultrametric and 
	#' to time calibrate the speciation nodes to fixed values
	#' 
	#' @param nhx A phylogenetic tree as a treeio::treedata object, with speciation 
	#' nodes annotated with a value of "N" in @data$D
	#' @param calibration_times A dataframe with two columns: age.min is the age 
	#' of a clade, and clade is the name of the clade
	#' @return A treeio::treedata object if successfully calibrated
	#' @export
	calibrate_tree = function ( nhx, calibration_times ) {

		tree = try( 
			ape::chronos( nhx@phylo, calibration=calibration, model="correlated" ) 
		)

@@

@@

@@

```{r prep_trees, echo=FALSE, cache=TRUE, message=F, warning=F}

@@

	# Make the trees ultrametric and calibrate the speciation nodes to 
	# specified times
	gene_trees_calibrated = 
		mclapply( 
			gene_trees_sub,
			calibrate_tree, 
			calibration_times=calibration_times,
			mc.cores=cores
		)

	# Parse the node ages from the internal @phylo object and store them with the 
	# corresponding rows in the @data object
	gene_trees_calibrated = 
		mclapply( 
			gene_trees_calibrated, 
			store_node_age, 
			mc.cores=cores 
		)

@@

```{r contrasts, cache=TRUE, echo=FALSE, message=F, warning=F}

	# Combine the @data slots across all trees. These have labels, node annotations,
	# expression data, contrast results, etc...
	all_nodes = 
		lapply( 
			gene_trees_pic, 
			function( nhx ){
				tags = nhx@data
				tags$gene = digest( nhx ) # Creates a  hash that is unique to each gene tree
				tags %<>% select( -( B ) ) # B has inconsistent types, remove it
				return( tags )
			}
		) %>%
		bind_rows()

@@

```{r show_result, echo=FALSE, cache=TRUE, message=F, warning=F}

@@

	# Check if the mean rank of duplication pics is greater than speciation pics
	wilcox_test_result = 
		wilcox.test( 
			nodes_contrast %>% filter( D=="Y" ) %>% .$pic, 
			nodes_contrast %>% filter( D=="N" ) %>% .$pic, 
			alternative="greater" 
		)

@@

Given the ortholog conjecture, we would expect to see larger contrasts following duplication events than speciation events. Instead, we find no such distribution (Figure XXKMRR). The Wilcox rank test does not reject the null hypothesis that the rate of evolution following duplications is the same as or less than the rate following speciation (p value = `r wilcox_test_result$p.value`). Our results therefore find no evidence for the ortholog conjecture in this system.

@@

	wilcox_filtered_age = 
		wilcox.test( 
			nodes_contrast_filtered_age %>% filter( D=="Y" ) %>% .$pic, 
			nodes_contrast_filtered_age %>% filter( D=="N" ) %>% .$pic, 
			alternative="greater" 
		)$p.value

@@

	wilcox_filtered_var = 
	wilcox.test( 
		nodes_contrast_filtered_var %>% filter( D=="Y" ) %>% .$pic, 
		nodes_contrast_filtered_var %>% filter( D=="N" ) %>% .$pic, 
		alternative="greater" 
	)$p.value

>>>>

	# These clade dates are from the KMRR code, they got them from http://timetree.org/
	calibration_times = data.frame(
		age.min =
			c( 20, 92, 167, 8, 42, 74, 296, 535, 104, 937, 29, 722, 441, 65, 15, 1215, 
				414, 371, 162, 25, 74, 77, 86 ), 
		clade = 
			c("Hominoidea", "Euarchontoglires", "Mammalia", "Homininae", 
					"Simiiformes", "Primates", "Amniota", "Vertebrata", "Eutheria", 
					"Bilateria", "Catarrhini", "Chordata", "Euteleostomi", 
					"Haplorrhini", "Hominidae", "Opisthokonta", "Sarcopterygii", 
					"Tetrapoda", "Theria", "Murinae", "Sciurognathi", "Rodentia", "Glires"
			),
		stringsAsFactors=FALSE
	)
```

```{r define_functions, echo=FALSE, message=F, warning=F}

@@

@@

	#' Adjust branch lengths of a phylogenetic tree to make it ultrametric and 
	#' to time calibrate the speciation nodes to fixed values
	#' 
	#' @param nhx A phylogenetic tree as a treeio::treedata object, with speciation 
	#' nodes annotated with a value of "N" in @data$D
	#' @param calibration_times A dataframe with two columns: age.min is the age 
	#' of a clade, and clade is the name of the clade
	#' @param ... Any additional arguments to pass to ape::chronos()
	#' @return A treeio::treedata object if successfully calibrated
	#' @export
	calibrate_tree = function ( nhx, calibration_times, ... ) {

		tree = try( 
			ape::chronos( nhx@phylo, calibration=calibration ) 
		)

@@

	#' Add phylogenetic independent contrasts (PIC) with the tree and 
	#' character data in a treeio::treedata object
	#' 
	#' @param gene_trees_calibrated A list of time calibrated phylogenetic trees 
	#' as treeio::treedata objects 
	#' @return A list of time phylogenetic trees with pics 
	#' as treeio::treedata objects
	#' @export
	add_pics_to_trees = function( gene_trees_calibrated ) {

		return( gene_trees_pic )
	}

	#' Add phylogenetic independent contrasts (PIC) with the tree and 
	#' character data in a treeio::treedata object
	#' 
	#' @param gene_trees_pic A list of phylogenetic trees with pics
	#' as treeio::treedata objects 
	#' @return A tibble with combined data, including phylogenetic independent 
	#' contrasts, for internal nodes of all trees
	#' @export
	summarize_contrasts = function( gene_trees_pic ) {
		# Combine the @data slots across all trees. These have labels, node annotations,
		# expression data, contrast results, etc...
		all_nodes = 
			lapply( 
				gene_trees_pic, 
				function( nhx ){
					tags = nhx@data
					tags$gene = digest( nhx ) # Creates a  hash that is unique to each gene tree
					tags %<>% select( -( B ) ) # B has inconsistent types, remove it
					return( tags )
				}
			) %>%
			bind_rows()

		return( nodes_contrast )
	}

	#' Calibrate a list of gene trees given a set of speciation node dates
	#' 
	#' @param gene_trees_sub A list of phylogenetic trees as treeio::treedata objects 
	#' with associated character data
	#' @param calibration_times A dataframe with two columns: age.min is the age 
	#' of a clade, and clade is the name of the clade
	#' @param ... Any additional arguments to pass to ape::chronos()
	#' @return A list of calibrated phylogenetic trees as treeio::treedata objects. May be 
	#' shorter than gene_trees_sub if some calibrations fail.
	#' @export
	calibrate_trees = function( gene_trees_sub, calibration_times, ... ) {
		# Make the trees ultrametric and calibrate the speciation nodes to 
		# specified times
		gene_trees_calibrated = 
			mclapply( 
				gene_trees_sub,
				calibrate_tree, 
				calibration_times=calibration_times,
				...,
				mc.cores=cores
			)

		# Parse the calibrated node ages from the internal @phylo object and store them with the 
		# corresponding rows in the @data object
		gene_trees_calibrated = 
			mclapply( 
				gene_trees_calibrated, 
				store_node_age, 
				mc.cores=cores 
			)

		return( gene_trees_calibrated )
	}

	#' Wilcoxon test of the ortholog conjecture, specifically that contrasts
	#' associated with duplication nodes are greater than those for 
	#' speciation nodes. A significant result indicates rejection of the null
	#' hypothesis that they are not greater.
	#' 
	#' @param nodes_contrast A tibble of independent contrasts
	#' @return Wilcoxon test p value
	#' @export
	wilcox_oc = function( nodes_contrast ) {

		p = wilcox.test( 
			nodes_contrast %>% filter( D=="Y" ) %>% .$pic, 
			nodes_contrast %>% filter( D=="N" ) %>% .$pic, 
			alternative="greater" 
		)$p.value

		return( p )
	}

@@

@@

@@

```{r calibrate_trees, echo=FALSE, cache=TRUE, message=F, warning=F}

	# Time calibrate all speciation nodes to the same years
	gene_trees_calibrated = calibrate_trees( gene_trees_sub, calibration_times, model="correlated" )

@@

```{r trees_to_contrasts, cache=TRUE, echo=FALSE, message=F, warning=F}

	# Calculate the contrasts from the trees and associated expression data
	gene_trees_pic = add_pics_to_trees( gene_trees_calibrated )

	# Collect all the contrasts in a single tibble
	nodes_contrast = summarize_contrasts( gene_trees_pic )

	# Test if the mean rank of duplication pics is greater than speciation pics
	wilcox_test_result = wilcox_oc( nodes_contrast )

```{r show_contrast_result, echo=FALSE, cache=TRUE, message=F, warning=F}

@@

@@

Given the ortholog conjecture, we would expect to see larger contrasts following duplication events than speciation events. Instead, we find no such distribution (Figure XXKMRR). The Wilcox rank test does not reject the null hypothesis that the rate of evolution following duplications is the same as or less than the rate following speciation (p value = `r wilcox_test_result`). Our results therefore find no evidence for the ortholog conjecture in this system.

@@

	wilcox_filtered_age = wilcox_oc( nodes_contrast_filtered_age )

@@

	wilcox_filtered_var = wilcox_oc( nodes_contrast_filtered_var )
